Language-Based Security for Low-Level MPC

### 2024-08-26 csf25/ctj.tex
checkpoint

For example, given a message send such as:
$$
\xassign{\mx{x}{1}}{\secret{x} \ftimes \flip{x}}{2}
$$
we would assign the following type to denote that $\mx{x}{1}$ is dependent on
both $\sx{x}{2}$ and $\rx{x}{2}$:
$$
\mx{x}{1} : \setit{\sx{x}{2}, \rx{x}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{x}{1}}{\secret{x} \fminus \flip{x}}{2}
$$
we want the type of $\mx{x}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{x}{2}$. However, we also
want to track the dependence of $\mx{x}{2}$ and $\rx{x}{2}$.
In short, we assign $\mx{x}{1}$ a ``ciphertext type'':
$$
\mx{x}{1} : \cty{\rx{x}{2}}{\sx{x}{2}}
$$
indicating that $\mx{x}{1}$ is an encryption of a value
dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client 1
does not subsequently receive messages that are dependent on $\rx{x}{2}$, the
value $\sx{x}{2}$ remains hidden.

Of course, there are other methods for encrypting values
in MPC protocols-- but we can observe that many are algebraically
equivalent to this fundamental schema. For example, in
Yao's Garbled Circuits (YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents
encrypted wire values as random samples, and the ``evaluator''
obtains the wire value for its input via OT. The evaluator shares 
the initial encrypted wire value of one of its input secrets as
a selection bit-- assuming that client
2 is the garbler and client 1 is the evaluator, we can
write this as:
$$
\xassign{\mx{x}{1}}{\mux{\secret{x}}{\neg\flip{x}}{\flip{x}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{x}{1}}{\OT{\secret{x}}{1}{\neg\flip{x}}{\flip{x}}}{2}}
   \toeq{\prog} \models \mx{x}{1} \eop \sx{x}{2} \fplus \rx{x}{2}
\end{mathpar}
This allows us to assign the same ciphertext type to $\mx{x}{1}$ as
above-- that is, $\cty{\rx{x}{2}}{\sx{x}{2}}$.

$$
\begin{array}{rcl@{\hspace{2mm}}r}
  t &::=& x \mid \cty{x}{T} \\
  \ty & \in & 2^{t}\\
  \Gamma &::=& \varnothing \mid \Gamma; x : \ty
\end{array} 
$$

\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
\end{definition}

\begin{mathpar}
  \inferrule[DepTy]
  {}
  {\eqj{\varnothing}{\eqs}{\phi}{\vars(\phi)}}

@@

  \inferrule[Seq]
            {\cpj{R_1}{\eqs}{\phi_1}{\Gamma_1}\\
             \cpj{R_2}{\eqs}{\phi_2}{\Gamma_2}}
            {\cpj{R_1;R_2}{\eqs}{\phi_1 \wedge \phi_2}{\Gamma_1;\Gamma_2}}
\end{mathpar}

\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}

\begin{mathpar}
  \inferrule
      {\cid \in C}
      {\leakj{\Gamma}{C}{\Gamma(\mx{w}{\cid})}}

@@

  \inferrule
      {\leakj{\Gamma}{C}{\setit{ \rx{w}{\cid} }} \\ \leakj{\Gamma}{C}{\setit{ \cty{\rx{w}{\cid}}{\ty} }} }
      {\leakj{\Gamma}{C}{\ty}}
\end{mathpar}

@@

\subsection{Examples}

@@

// m[y]@1 == s[y]@1 + -r[y]@2
// m[y]@1 : { c(r[y]@2, { s[y]@1 }) } 
\end{verbatimtab}

$$
\select{x}{y}{z}
$$

>>>>

For example, given a message send such as:
$$
\xassign{\mx{x}{1}}{\secret{x} \ftimes \flip{x}}{2}
$$
we would assign the following type to denote that $\mx{x}{1}$ is dependent on
both $\sx{x}{2}$ and $\rx{x}{2}$:
$$
\mx{x}{1} : \setit{\sx{x}{2}, \rx{x}{2}}
$$
But in the case of 2-party reconstructive sharing:
$$
\xassign{\mx{x}{1}}{\secret{x} \fminus \flip{x}}{2}
$$
we want the type of $\mx{x}{1}$ to be an ``encrypted'' type
that is by itself independent of $\sx{x}{2}$. However, we also
want to track the dependence of $\mx{x}{2}$ and $\rx{x}{2}$.
In short, we assign $\mx{x}{1}$ a ``ciphertext type'':
$$
\mx{x}{1} : \cty{\rx{x}{2}}{\setit{\sx{x}{2}}}
$$
indicating syntactically that $\mx{x}{1}$ is an encryption of a
value dependent on $\sx{x}{2}$ with $\rx{x}{2}$ and as long as client
1 does not subsequently receive messages that are dependent on
$\rx{x}{2}$, the value $\sx{x}{2}$ remains hidden.

Of course, there are other methods for encrypting values in MPC
protocols-- but we can observe that many are algebraically equivalent
to this fundamental schema. For example, in Yao's Garbled Circuits
(YGC) in $\mathbb{F}_{2}$, the ``garbler'' represents encrypted wire
values as a random sample (for 1) or its negation (for 0), and the
``evaluator'' obtains the wire value for its input via OT. The
evaluator shares the initial encrypted wire value of one of its input
secrets as a selection bit-- assuming that client 2 is the garbler and
client 1 is the evaluator, we can write this as:
$$
\xassign{\mx{x}{1}}{\mux{\secret{x}}{\neg\flip{x}}{\flip{x}}}{2}
$$
where for all $\be_1,\be_2,\be_3$:
$$
\mux{\be_1}{\be_2}{\be_3} \defeq (\neg\be_1 \ftimes \be_2) \fplus (\be_1 \ftimes \be_3)
$$
and letting this protocol be $\prog$ the following is valid:
\begin{mathpar}
%\toeq{\xassign{\mx{x}{1}}{\OT{\secret{x}}{1}{\neg\flip{x}}{\flip{x}}}{2}}
   \toeq{\prog} \models \mx{x}{1} \eop \neg\sx{x}{2} \fplus \rx{x}{2}
\end{mathpar}
This allows us to assign the same ciphertext type to $\mx{x}{1}$ as
above-- that is, $\cty{\rx{x}{2}}{\setit{\sx{x}{2}}}$. Leveraging the
equivalence this way has been previously well-studied
\cite{barthe2019probabilistic}.  Since this is a one-time-pad scheme,
it is also important that we ensure that samples are used at most once
for encryption. As in previous work \cite{darais2019language} we can
use a form of type linearity to this end.

The syntax of types $\ty$ and type environments $\Gamma$ are presented
in Figure \ref{fig-cpj}\cnote{need to fix set notations}. We also
define type judgements for expressions and programs by direct
interpretation as terms $\phi$ and constraints $\eqs$ respectively,
via the encoding $\toeq{cdot}$. The \TirName{Encode} rule captures
the use ``fresh'' samples for encryption discussed above. Derivations
enforce linearity by requiring that variables used for encryption
are added to judgements, and guaranteed to not be used elsewhere
via a form of union that ensures disjointness of combined sets. 
\begin{definition}
  $R_1;R_2 = R_1 \cup R_2$ iff $R_1 \cap R_2 = \varnothing$.
\end{definition}
The \TirName{DepTy} rule handles all other syntactic forms and just
records the dependencies on variables in the term.  The \TirName{Send}
rule records the types of messages, reveals, and outputs in
enviroments, and the \TirName{Seq} rule combines the types of
sequenced programs.  Given this we define validity of program type
judgements as follows. \cnote{Define $\eqspre$.}
\begin{definition}
  Given preprocessing predicate $\eqspre$ and protocol $\prog$ we say
  $\cpj{R}{\eqs}{\eqspre \wedge \toeq{\prog}}{\Gamma}$ is \emph{valid} iff it is derivable and
  $\eqspre \wedge \toeq{\prog} \models \eqs$.
\end{definition}
It is important to note that this presentation of derivations is
logical, not algorithmic-- in particular it is not syntax-directed
due to the \TirName{Encode} rule.

\begin{fpfig}[t]{Syntax and Derivation Rules for $\minicat$ Confidentiality Types}{fig-cpj}
\small{
$$
\begin{array}{rcl@{\hspace{3mm}}l}
  t &::=& x \mid \cty{x}{T} \\
  \ty & \in & 2^{t} & \gdesc{types}\\
  \Gamma &::=& \varnothing \mid \Gamma; x : \ty & \gdesc{type environments}
\end{array} 
$$
\medskip
\begin{mathpar}
  \inferrule[DepTy]
  {}
  {\eqj{\varnothing}{\eqs}{\phi}{\vars(\phi)}}

@@

  \inferrule[Seq]
            {\cpj{R_1}{\eqs}{\phi_1}{\Gamma_1}\\
             \cpj{R_2}{\eqs}{\phi_2}{\Gamma_2}}
            {\cpj{R_1;R_2}{\eqs}{\phi_1 \wedge \phi_2}{\Gamma_1;\Gamma_2}}
\end{mathpar}
}
\end{fpfig}

\begin{fpfig}[t]{Dependencies in Corrupt Views: Derivation Rules}{fig-leakj}
\small{
\begin{mathpar}
  \inferrule
      {\cid \in C}
      {\leakj{\Gamma}{C}{\Gamma(\mx{w}{\cid})}}

@@

  \inferrule
      {\leakj{\Gamma}{C}{\setit{ \rx{w}{\cid} }} \\ \leakj{\Gamma}{C}{\setit{ \cty{\rx{w}{\cid}}{\ty} }} }
      {\leakj{\Gamma}{C}{\ty}}
\end{mathpar}
}
\end{fpfig}

@@

\begin{comment}
\subsection{Examples}

@@

// m[y]@1 == s[y]@1 + -r[y]@2
// m[y]@1 : { c(r[y]@2, { s[y]@1 }) } 
\end{verbatimtab}
\end{comment}
